---
title: "Final Project Proposal"
author: "Ian Beller (1006837620)"
date: "DUE: March 21st 2022"
output: pdf_document
urlcolor: blue
editor_options: 
  chunk_output_type: inline
---

```{r message=FALSE}
library(tidyverse)
library(latex2exp)
library(gridExtra)
library(dplyr)

all.beijing <- read.csv('./LSTM-Multivariate_pollution.csv')
glimpse(all.beijing)
```

```{r}
drops <- c("snow", 'press', 'rain', 'wnd_dir')
beijing <- all.beijing[ , !(names(all.beijing) %in% drops)]
glimpse(beijing)

```



```{r}
sum(is.na(beijing))
```

```{r}
beijing_summ <- data.frame(do.call(cbind, lapply(beijing, summary)))
names(beijing_summ)[names(beijing_summ) == "temp"] <- "temperature"
names(beijing_summ)[names(beijing_summ) == "wnd_spd"] <- "wind_speed"
beijing_summ[1, 1] = beijing$date[[1]]
beijing_summ[2, 1] = beijing$date[[10950]]
beijing_summ[3, 1] = beijing$date[[21900]]
beijing_summ[4, 1] = '-'
beijing_summ[5, 1] = beijing$date[[32850]]
beijing_summ[6, 1] = beijing$date[[43800]]
beijing_summ[nrow(beijing_summ) + 1,] = c('-', IQR(beijing$pollution),
                                          IQR(beijing$temp), 
                                          IQR(beijing$wnd_spd),
                                          IQR(beijing$dew))
row.names(beijing_summ)[7] <- 'Interquartile Range'
beijing_summ[nrow(beijing_summ) + 1,] = c('-', sd(beijing$pollution), 
                                          sd(beijing$temp), 
                                          sd(beijing$wnd_spd), 
                                          sd(beijing$dew))
row.names(beijing_summ)[8] <- 'Standard Deveation'
beijing_summ[nrow(beijing_summ) + 1,] = c('-', var(beijing$pollution), 
                                          var(beijing$temp), 
                                          var(beijing$wnd_spd), 
                                          var(beijing$dew))
row.names(beijing_summ)[9] <- 'Variance'
beijing_summ[nrow(beijing_summ) + 1,] = c('-', mad(beijing$pollution), 
                                          mad(beijing$temp), 
                                          mad(beijing$wnd_spd), 
                                          mad(beijing$dew))
row.names(beijing_summ)[10] <- 'Median Absolute deviation'
beijing_summ
```

```{r}
beijing$date <- as.Date(beijing$date)
sum.dates <- (aggregate(beijing["pollution"], by=beijing["date"], sum))[1]
sum.polu <- (aggregate(beijing["pollution"], by=beijing["date"], sum))[2]
sum.dew <- (aggregate(beijing["dew"], by=beijing["date"], sum))[2]
sum.temp <- (aggregate(beijing["temp"], by=beijing["date"], sum))[2]
sum.w_s <- (aggregate(beijing["wnd_spd"], by=beijing["date"], sum))[2]

avg.beijing <- data.frame(sum.dates, sum.polu, sum.temp, sum.dew, sum.w_s)
avg.beijing$pollution <- as.numeric(as.character(avg.beijing$pollution)) / 24
avg.beijing$temp <- as.numeric(as.character(avg.beijing$temp)) / 24
avg.beijing$dew <- as.numeric(as.character(avg.beijing$dew)) / 24
avg.beijing$wnd_spd <- as.numeric(as.character(avg.beijing$wnd_spd)) / 24
avg.beijing$date <- as.character(avg.beijing$date)
```


```{r}
glimpse(avg.beijing)
```

```{r}
avg.beijing_summ <- data.frame(do.call(cbind, lapply(avg.beijing, summary)))
names(avg.beijing_summ)[names(avg.beijing_summ) == "temp"] <- "temperature"
names(avg.beijing_summ)[names(avg.beijing_summ) == "wnd_spd"] <- "wind_speed"
avg.beijing_summ[1, 1] = avg.beijing$date[[1]]
avg.beijing_summ[2, 1] = avg.beijing$date[[456]]
avg.beijing_summ[3, 1] = avg.beijing$date[[912]]
avg.beijing_summ[4, 1] = '-'
avg.beijing_summ[5, 1] = avg.beijing$date[[1369]]
avg.beijing_summ[6, 1] = avg.beijing$date[[1825]]
avg.beijing_summ[nrow(avg.beijing_summ) + 1,] = c('-', 
                                                  IQR(avg.beijing$pollution), 
                                                  IQR(avg.beijing$temp), 
                                                  IQR(avg.beijing$wnd_spd), 
                                                  IQR(avg.beijing$dew))
row.names(avg.beijing_summ)[7] <- 'Interquartile Range'
avg.beijing_summ[nrow(avg.beijing_summ) + 1,] = c('-', 
                                                  sd(avg.beijing$pollution), 
                                                  sd(avg.beijing$temp),
                                                  sd(avg.beijing$wnd_spd), 
                                                  sd(avg.beijing$dew))
row.names(avg.beijing_summ)[8] <- 'Standard Deveation'
avg.beijing_summ[nrow(avg.beijing_summ) + 1,] = c('-', 
                                                  var(avg.beijing$pollution),
                                                  var(avg.beijing$temp), 
                                                  var(avg.beijing$wnd_spd), 
                                                  var(avg.beijing$dew))
row.names(avg.beijing_summ)[9] <- 'Variance'
avg.beijing_summ[nrow(avg.beijing_summ) + 1,] = c('-', 
                                                  mad(avg.beijing$pollution), 
                                                  mad(avg.beijing$temp), 
                                                  mad(avg.beijing$wnd_spd), 
                                                  mad(avg.beijing$dew))
row.names(avg.beijing_summ)[10] <- 'Median Absolute deviation'
avg.beijing_summ
```



```{r fig.dim=c(8, 5), fig.align="center"}
boxpollution <- ggplot(beijing)+
  geom_boxplot(aes(x=pollution), outlier.colour="red", outlier.shape=20,
                outlier.size=2)  + 
  stat_boxplot(geom ='errorbar',  aes(x=pollution), width=0.6) +
  labs(x='Pollution level (in levels of PM2.5 concentration)', 
       title='Pollution levels per hour in Beijing between 2010-2015', 
       subtitle='43800 entries registered by the U.S. embassy in Beijing') +
  scale_y_continuous(labels=NULL, breaks=NULL) +
  theme_gray()

boxtemp <- ggplot(beijing)+
  geom_boxplot(aes(x=temp), outlier.colour="red", outlier.shape=20,
                outlier.size=2)  + 
  stat_boxplot(geom ='errorbar',  aes(x=temp), width=0.6) +
  labs(x='Temperature (in Celsius)', 
       title='Temperature per hour in Beijing between 2010-2015', 
       subtitle='43800 entries registered by the U.S. embassy in Beijing') +
  scale_y_continuous(labels=NULL, breaks=NULL) +
  theme_gray()

boxwspd <- ggplot(beijing)+
  geom_boxplot(aes(x=wnd_spd), outlier.colour="red", outlier.shape=20,
                outlier.size=2)  + 
  stat_boxplot(geom ='errorbar',  aes(x=wnd_spd), width=0.6) +
  labs(x='wind speed (in knots)', 
       title='wind speed in Beijing between 2010-2015', 
       subtitle='43800 entries registered by the U.S. embassy in Beijing') +
  scale_y_continuous(labels=NULL, breaks=NULL) +
  theme_gray()

grid.arrange(grobs = list(boxpollution, boxtemp, boxwspd), nrows=3)
```



```{r warning=FALSE, fig.align="center", fig.dim=c(12, 9)}

pollution_qq <- avg.beijing %>% 
  ggplot(aes(sample= pollution)) +
  geom_qq()+
  geom_qq_line() +
  labs(x = 'N(0, 1) percentiles',
       y = '2.5PM Percentiles', 
       title = 'Normal QQ plot of Beijing pollution \nlevels from 2010 to 2015') + 
  theme_light()

temp_qq <- avg.beijing %>% 
  ggplot(aes(sample= temp)) +
  geom_qq()+
  geom_qq_line() +
  labs(x = 'N(0, 1) percentiles',
       y = 'Temperature Percentiles', 
       title = 'Normal QQ plot of Beijing temperature \nlevels (in ªC) from 2010 to 2015') + 
  theme_light()

dew_qq <- avg.beijing %>% 
  ggplot(aes(sample= dew)) +
  geom_qq()+
  geom_qq_line() +
  labs(x = 'N(0, 1) percentiles',
       y = 'Dew points Percentiles', 
       title = 'Normal QQ plot of Beijing dew \npoints from 2010 to 2015') + 
  theme_light()

ws_qq <- avg.beijing %>% 
  ggplot(aes(sample= wnd_spd)) +
  geom_qq()+
  geom_qq_line() +
  labs(x = 'N(0, 1) percentiles',
       y = 'Wind speed Percentiles', 
       title = 'Normal QQ plot of Beijing wind \nspeed from 2010 to 2015') + 
  theme_light()
grid.arrange(grobs = list(pollution_qq, temp_qq, dew_qq, ws_qq), nrows=2, ncols= 2)
```





```{r}
model <- lm(pollution ~ temp + wnd_spd + dew, data = avg.beijing)
summary(model)
```
```{r}
b0 <- 181.4200
se_b0 <- 3.8240
b1 <- -7.3376
se_b1 <- 0.3213
b2 <- -0.2832
se_b2 <- 0.0403
b3 <- 5.8957
se_b3 <- 0.2714

print('b0')
lb <- b0 - qt(0.025, 1821) * se_b0
ub <- b0 + qt(0.025, 1821) * se_b0
quantile(c(ub, lb), probs= c(0.975, 0.025))

print('b1')
lb <- b1 - qt(0.025, 1821) * se_b1
ub <- b1 + qt(0.025, 1821) * se_b1
quantile(c(ub, lb), probs= c(0.975, 0.025))

print('b2')
lb <- b2 - qt(0.025, 1821) * se_b2
ub <- b2 + qt(0.025, 1821) * se_b2
quantile(c(ub, lb), probs= c(0.975, 0.025))

print('b3')
lb <- b3 - qt(0.025, 1821) * se_b3
ub <- b3 + qt(0.025, 1821) * se_b3
quantile(c(ub, lb), probs= c(0.975, 0.025))
```


```{r}
avg.beijing$month<- format(as.Date(avg.beijing$date), "%m")

X <- avg.beijing[avg.beijing$month <= '05',]
spring <- X[X$month >= '04',]

X <- avg.beijing[avg.beijing$month <= '08',]
summer <- X[X$month >= '06',]

X <- avg.beijing[avg.beijing$month <= '10',]
autumn  <- X[X$month >= '09',]

Y <- avg.beijing[avg.beijing$month > '10',]
X <- avg.beijing[avg.beijing$month <= '03',]
winter  <- rbind(X, Y)


glimpse(spring)
glimpse(summer)
glimpse(autumn)
glimpse(winter)
```


```{r}
model_spring <- lm(pollution ~ temp + wnd_spd + dew, data = spring)
summary(model_spring)
```
```{r}
b0 <- 127.7242
se_b0 <- 10.0182
b1 <- -3.2278
se_b1 <- 0.5734
b2 <- -0.1340
se_b2 <- 0.0633
b3 <- 4.0631
se_b3 <- 0.3782

print('b0')
lb <- b0 - qt(0.025, 301) * se_b0
ub <- b0 + qt(0.025, 301) * se_b0
quantile(c(ub, lb), probs= c(0.975, 0.025))

print('b1')
lb <- b1 - qt(0.025, 301) * se_b1
ub <- b1 + qt(0.025, 301) * se_b1
quantile(c(ub, lb), probs= c(0.975, 0.025))

print('b2')
lb <- b2 - qt(0.025, 301) * se_b2
ub <- b2 + qt(0.025, 301) * se_b2
quantile(c(ub, lb), probs= c(0.975, 0.025))

print('b3')
lb <- b3 - qt(0.025, 301) * se_b3
ub <- b3 + qt(0.025, 301) * se_b3
quantile(c(ub, lb), probs= c(0.975, 0.025))

```

```{r}
model_winter <- lm(pollution ~ temp + wnd_spd + dew, data = winter)
summary(model_winter)
```
```{r}
b0 <- 222.1861
se_b0 <- 5.8442
b1 <- -6.4398
se_b1 <- 0.5755
b2 <- -0.2547
se_b2 <- 0.0495
b3 <- 9.0277
se_b3 <- 0.4614

print('b0')
lb <- b0 - qt(0.025, 751) * se_b0
ub <- b0 + qt(0.025, 751) * se_b0
quantile(c(ub, lb), probs= c(0.975, 0.025))

print('b1')
lb <- b1 - qt(0.025, 751) * se_b1
ub <- b1 + qt(0.025, 751) * se_b1
quantile(c(ub, lb), probs= c(0.975, 0.025))

print('b2')
lb <- b2 - qt(0.025, 751) * se_b2
ub <- b2 + qt(0.025, 751) * se_b2
quantile(c(ub, lb), probs= c(0.975, 0.025))

print('b3')
lb <- b3 - qt(0.025, 751) * se_b3
ub <- b3 + qt(0.025, 751) * se_b3
quantile(c(ub, lb), probs= c(0.975, 0.025))
```


```{r}
model_autumn <- lm(pollution ~ temp + wnd_spd + dew, data = autumn)
summary(model_autumn)
```
```{r}
model_autumn <- lm(pollution ~ temp + dew, data = autumn)
summary(model_autumn)
```
```{r}
b0 <- 211.105
se_b0 <- 17.455
b1 <- -13.134
se_b1 <- 1.385
b3 <- 11.536
se_b3 <- 0.953

print('b0')
lb <- b0 - qt(0.025, 302) * se_b0
ub <- b0 + qt(0.025, 302) * se_b0
quantile(c(ub, lb), probs= c(0.975, 0.025))

print('b1')
lb <- b1 - qt(0.025, 302) * se_b1
ub <- b1 + qt(0.025, 302) * se_b1
quantile(c(ub, lb), probs= c(0.975, 0.025))

print('b3')
lb <- b3 - qt(0.025, 302) * se_b3
ub <- b3 + qt(0.025, 302) * se_b3
quantile(c(ub, lb), probs= c(0.975, 0.025))
```


```{r}
model_summer <- lm(pollution ~ temp + dew + wnd_spd , data = summer)
summary(model_summer)
```
We notice that wnd_spd parameter is statistically insignificant since it has a
p-value above 0.05. We also see that the intercept value is also statistically 
insignificant and this skews with our model since we do not have enough evidence
to reject the null hypothesis for our variables which means that we do not have
enough evidence to assume our regression line is not horizontal/ there is no 
relationship. Hence to have the model with the most precision when estimating 
pollution levels in summer, we will disregard the parameter wnd_spd. Hence our 
statistical summary for our new model is 
```{r}
model_summer <- lm(pollution ~ temp + dew, data = summer)
summary(model_summer)
```
In the new regression model, we have the same issue as our regression model but
now the temperature is statistically insignificant which keeps the intercept 
also statistically insignificant. Hence we will re-run another regression model 
but now disregard temperature as well.

```{r}
model_summer <- lm(pollution ~dew, data = summer)
summary(model_summer)
```
```{r}
b0 <- -47.258
se_b0 <- 11.318
b3 <- 7.086
se_b3 <- 0.594

print('b0')
lb <- b0 - qt(0.025, 458) * se_b0
ub <- b0 + qt(0.025, 458) * se_b0
quantile(c(ub, lb), probs= c(0.975, 0.025))

print('b3')
lb <- b3 - qt(0.025, 458) * se_b3
ub <- b3 + qt(0.025, 458) * se_b3
quantile(c(ub, lb), probs= c(0.975, 0.025))
```


